{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-10 thesis/tm.tex
more tm tests

\begin{table*}[t]
	\begin{center}
		\footnotesize
	\begin{tabular}{cc||r|r|r||r|r|r|r}
			&	&\multicolumn{3}{c||}{\bf Quicksand mode}&\multicolumn{4}{c}{{\bf Maximal state space mode} ({\tt -M})} \\
		\bf buggy test	& \bf params&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
		\hline
		\hline
		{\tt htm1}
			& 2,1	&\cpu{  26.48}&\wtm{6.48}&\ints{5}& \cpu{*8.21}& \wtm{*5.78}& \ints{5	}& \ETA{12} \\
		(assertion)
			& 2,2	&\cpu{  45.23}&\wtm{8.28}&\ints{9}& \cpu{*8.42}& \wtm{*5.97}& \ints{9	}& \ETA{102} \\
			& 2,3	&\cpu{ 53.88}&\wtm{9.92}&\ints{17}& \cpu{*8.80}& \wtm{*6.31}& \ints{17	}& \ETA{819} \\
			& 2,4	&\cpu{ 90.24}&\wtm{14.63}&\ints{33}& \cpu{*9.72}& \wtm{*7.19}& \ints{33	}& \ETA{6553} \\
		\cline{2-9}
			& 3,1	&\cpu{ 39.75}&\wtm{8.21}&\ints{5}& \cpu{*8.15}& \wtm{*5.71}& \ints{5	}& \ETA{76} \\
			& 3,2	&\cpu{ 46.91}&\wtm{8.40}&\ints{9}& \cpu{*8.37}& \wtm{*5.91}& \ints{9	}& \ETA{3686} \\
			& 3,3	&\cpu{ 41.39}&\wtm{8.50}&\ints{17}& \cpu{*8.71}& \wtm{*6.26}& \ints{17	}& \ETA{176947} \\
			& 3,4	&\cpu{ 92.98}&\wtm{15.19}&\ints{33}& \cpu{*10.36}& \wtm{*7.15}& \ints{33	}& \ETA{8493465} \\
		\cline{2-9}
			& 4,1	&\cpu{ 45.13}&\wtm{8.56}&\ints{5}& \cpu{*7.77}& \wtm{*5.30}& \ints{5	}& \ETA{460} \\
			& 4,2	&\cpu{64.99}&\wtm{11.33}&\ints{9}& \cpu{*8.32}& \wtm{*5.59}& \ints{9	}& \ETA{132710} \\
		\hline
		{\tt swapbug}
			& 2,1	&\cpu{104.37}&\wtm{17.00}&\ints{*6}& \cpu{*43.76}& \wtm{*12.51}& \ints{33	}& \ETA{73} \\
		(deadlock)
			& 2,2	&\cpu{ 66.47}&\wtm{*12.42}&\ints{*38}& \cpu{*45.22}& \wtm{15.30}& \ints{85	}& \ETA{860} \\
			& 2,3	&\cpu{ 99.41}&\wtm{*15.92}&\ints{*48}& \cpu{*53.57}& \wtm{21.34}& \ints{217	}& \ETA{9120} \\
			& 2,4	&\cpu{125.23}&\wtm{*19.46}&\ints{*68}& \cpu{*71.01}& \wtm{38.86}& \ints{537	}& \ETA{91239} \\
		\cline{2-9}
			& 3,1	&\cpu{313.36}&\wtm{*44.50}&\ints{*34}&\cpu{*78.23}& \wtm{60.49}& \ints{1016	}& \ETA{3543} \\
			& 3,2	&\cpu{*6500.36}&\wtm{*816.27}&\ints{*262}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{2161715} \\
		\hline
		{\tt avl\_insert}
			& 2,2	&\cpu{1355.05}&\wtm{174.09}&\ints{*29}	&\cpu{*93.28}&\wtm{*30.37}&\ints{79}&\ETA{157286} \\
		(segfault+)
			& 2,3	&\cpu{402.91}&\wtm{*54.60}&\ints{*33}&\cpu{*228.16}&\wtm{125.89}&\ints{835}&\ETA{13664151} \\
			% TODO: these 2,3 ones alone are new + ok
			% 33 interleavings; 1 preemptions; job time 12s; pldi time 179368977; new-fixed pldi cputime 1403360474
			% 835 interleavings; 1 preemptions; job time 1m 42s; pldi time 135292827; new-fixed pldi cputime 235664920
			& 2,4	&\ETA{--}&\ETA{>1h}&\ETA{--}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{79020639} \\
			% this quicksand jerk required ulimit -n higher than 1024
			%
			% the jerk didn't even find the bug anymore, after dpor-preferred tid
			% heres the stats from how it did before
			% & 2,4	&\cpu{*7598.19}&\wtm{*970.67}&\ints{*2633}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{79020639} \\
			% anyway this was the maximal
			%
		\cline{2-9}
			& 3,1	&\cpu{*86.43}&\wtm{*14.10}&\ints{*23}	&\cpu{91.97}&\wtm{24.80}&\ints{53}&\ETA{343250} \\
			& 3,2	&\cpu{*2177.86}&\wtm{*287.66}&\ints{*169}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{425939133} \\
			% TODO
			\multicolumn{9}{c}{everything above this line needs rerun for dpor-preferred tid + thr\_bottom1} \\

		\hline
		{\tt lock\_fast}
		%({\tt spinlock})
			& 2,1	&\cpu{18.33}&\wtm{5.19}&\ints{2}&\cpu{*3.12}&\wtm{*3.12}&\ints{2}& \ETA{4} \\
			% 2 interleavings; 0 preemptions; job time 4s; pldi time 5191170; new-fixed pldi cputime 18328224
			% 2 interleavings; 0 preemptions; job time 2s; pldi time 3119081; new-fixed pldi cputime 3118912
		(perf)
			& 9,9	&\cpu{22.43}&\wtm{6.24}&\ints{2}&\cpu{*4.71}&\wtm{*4.71}&\ints{2}& \ETA{inf} \\
			% 2 interleavings; 0 preemptions; job time 3s; pldi time 6238515; new-fixed pldi cputime 22431755
			% 2 interleavings; 0 preemptions; job time 3s; pldi time 4712012; new-fixed pldi cputime 4711884
	\end{tabular}
	\end{center}
	\caption{Landslide's bug-finding performance on various test configurations.
		Iterative Deepening (\sect{\ref{sec:quicksand-id}}), optimized for fast bug-finding,
		is compared against Maximal State Space mode (\sect{\ref{sec:landslide-quicksand-options}}),
		optimized for fast verification.
		For each, I list the CPU-time and wall-clock time elapsed,
		plus the number of interleavings tested in the ultimately buggy state space
		until the bug was found.
		* marks the winning measurements between each series.
		Lastly, state space estimation (\sect{\ref{sec:landslide-estimate}})
		%though approximate at best,
		confers a sense of the exponential explosion.
	}
	\label{tab:tm-buges}
\end{table*}

@@

In addition to the bugs I intentionally wrote in {\tt htm1} and {\tt swapbug},
%to my pleasant surprise % only the former of these is a surprise now
Landslide also found two bugs in the ``real-world'' transactional algorithms I tested.
\begin{itemize}
	\item
	{\bf Atomicity violation.}
	{\tt avl\_insert} with any parameters higher than (2,1)
	exposed a previously-unknown bug in the transactional AVL tree.
	Figure~\ref{fig:avlbug} shows the root cause, essentially the {\tt htm1} bug in disguise.
	% landslide-trace-1521093835.18927.html
	This manifested
	alternately
	as a segfault (for test parameters (2,2) and (3,1))
	and
	% landslide-trace-1533274266.417257.html
	as a consistency-check assertion failure (for test parameters (2,3)).
	The presence of {\tt while (\_retry);} makes the necessary preemption window extremely small
	(between it and {\tt \_xbegin()}),
	making the bug extremely unlikely to manifest under stress testing,
	but Landslide is blind to such matters of chance.

@@

@@

			& 2,2 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \cpu{1285.06}	& \ints{21684} \\

@@

Firstly, the middle column of Table~\ref{tab:tm-verifs}
shows the impact of Landslide's experimental retry set reduction (\sect{\ref{sec:tm-retrysets}}).
Despite its conservative implementation,
it provides roughly 2-6x reduction in most tests, with up to 20x in extreme cases.
The biggest win is apparent in {\tt fig63},
the test contrived to induce as much reduction as possible
using transactional paths that conflict only with aborts and not with each other, and vice versa.
In {\tt fig63}(2,1), corresponding to Figure~\ref{fig:retry-set-example},
retry-set-enabled DPOR correctly prunes down to the optimal 6 interleavings,
while the baseline treats it identically to the fully-conflicting {\tt counter}(2,1).
Figure~\ref{fig:retry-reduction} depicts the difference between the state spaces explored by the two approaches.

>>>>

\begin{table*}[t]
	\begin{tabular}{cc||r|r|r||r|r|r|r}
			&	&\multicolumn{3}{c||}{\bf Quicksand mode}&\multicolumn{4}{c}{{\bf Maximal state space mode} ({\tt -M})} \\
		\bf buggy test	& \bf params&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
		{\tt htm1}
			& 2,1	&\cpu{  26.48}&\wtm{6.48}&\ints{5}& \cpu{*8.21}& \wtm{*5.78}& \ints{5	}& \ETA{12} \\
		(assertion)
			& 2,2	&\cpu{  45.23}&\wtm{8.28}&\ints{9}& \cpu{*8.42}& \wtm{*5.97}& \ints{9	}& \ETA{102} \\
			& 2,3	&\cpu{ 53.88}&\wtm{9.92}&\ints{17}& \cpu{*8.80}& \wtm{*6.31}& \ints{17	}& \ETA{819} \\
			& 2,4	&\cpu{ 90.24}&\wtm{14.63}&\ints{33}& \cpu{*9.72}& \wtm{*7.19}& \ints{33	}& \ETA{6553} \\
		\cline{2-9}
			& 3,1	&\cpu{ 39.75}&\wtm{8.21}&\ints{5}& \cpu{*8.15}& \wtm{*5.71}& \ints{5	}& \ETA{76} \\
			& 3,2	&\cpu{ 46.91}&\wtm{8.40}&\ints{9}& \cpu{*8.37}& \wtm{*5.91}& \ints{9	}& \ETA{3686} \\
			& 3,3	&\cpu{ 41.39}&\wtm{8.50}&\ints{17}& \cpu{*8.71}& \wtm{*6.26}& \ints{17	}& \ETA{176947} \\
			& 3,4	&\cpu{ 92.98}&\wtm{15.19}&\ints{33}& \cpu{*10.36}& \wtm{*7.15}& \ints{33	}& \ETA{8493465} \\
		\cline{2-9}
			& 4,1	&\cpu{ 45.13}&\wtm{8.56}&\ints{5}& \cpu{*7.77}& \wtm{*5.30}& \ints{5	}& \ETA{460} \\
			& 4,2	&\cpu{64.99}&\wtm{11.33}&\ints{9}& \cpu{*8.32}& \wtm{*5.59}& \ints{9	}& \ETA{132710} \\
		{\tt swapbug}
			& 2,1	&\cpu{104.37}&\wtm{17.00}&\ints{*6}& \cpu{*43.76}& \wtm{*12.51}& \ints{33	}& \ETA{73} \\
		(deadlock)
			& 2,2	&\cpu{ 66.47}&\wtm{*12.42}&\ints{*38}& \cpu{*45.22}& \wtm{15.30}& \ints{85	}& \ETA{860} \\
			& 2,3	&\cpu{ 99.41}&\wtm{*15.92}&\ints{*48}& \cpu{*53.57}& \wtm{21.34}& \ints{217	}& \ETA{9120} \\
			& 2,4	&\cpu{125.23}&\wtm{*19.46}&\ints{*68}& \cpu{*71.01}& \wtm{38.86}& \ints{537	}& \ETA{91239} \\
		\cline{2-9}
			& 3,1	&\cpu{313.36}&\wtm{*44.50}&\ints{*34}&\cpu{*78.23}& \wtm{60.49}& \ints{1016	}& \ETA{3543} \\
			& 3,2	&\cpu{*6500.36}&\wtm{*816.27}&\ints{*262}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{2161715} \\
			\multicolumn{9}{c}{everything above this line needs rerun for dpor-preferred tid + thr\_bottom1} \\
			& 2,2	&\cpu{2494.77}&\wtm{315.46}&\ints{*29}	&\cpu{*95.53}&\wtm{*30.56}&\ints{79}&\ETA{158505} \\
			% 79 interleavings; 1 preemptions; job time 9s; pldi time 30558236; new-fixed pldi cputime 95530430
			% 29 interleavings; 1 preemptions; job time 14s; pldi time 315461192; new-fixed pldi cputime 2494772471
		(segfault+)
			& 2,3	&\cpu{308.10}&\wtm{*43.95}&\ints{*33}&\cpu{*249.42}&\wtm{144.33}&\ints{835}&\ETA{13664203} \\
			% 835 interleavings; 1 preemptions; job time 1m 50s; pldi time 144332259; new-fixed pldi cputime 249418666
			% 33 interleavings; 1 preemptions; job time 10s; pldi time 43954391; new-fixed pldi cputime 308100075
			& 2,4	&\cpu{*2979.29}&\wtm{*390.34}&\ints{*1457}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{61882736} \\
			% 1457 interleavings; 4 preemptions; job time 3m 11s; pldi time 390342652; new-fixed pldi cputime 2979291067
			% maximal: progress: 22854/61882736 brs (0.036931%), ETA 9y 307d 18h 4m 41s (elapsed 51m 54s)
			% update for below, post thr-bottom1 fix - ok it FABed after all;
			% not sure if higher ulimit still necessary but i still used it to be safe
			% (old comment:) this quicksand jerk required ulimit -n higher than 1024
			%
			% the jerk didn't even find the bug anymore, after dpor-preferred tid
			% heres the stats from how it did before (older comment:)
			% & 2,4	&\cpu{*7598.19}&\wtm{*970.67}&\ints{*2633}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{79020639} \\
		\cline{2-9}
			& 3,1	&\cpu{*87.10}&\wtm{*14.81}&\ints{*14}	&\cpu{94.08}&\wtm{23.60}&\ints{24}&\ETA{207575} \\
			% 14 interleavings tested; 3 preemptions; job time 8s; pldi time 14807247; new-fixed pldi cputime 87102600
			% 24 interleavings tested; 3 preemptions; job time 5s; pldi time 23600679; new-fixed pldi cputime 94082790
			& 3,2	&\cpu{*3672.84}&\wtm{*475.03}&\ints{*145}&\ETA{--}&\ETA{>1h}&\ETA{--}& \ETA{1635075071} \\
			% maximal: progress: 6571/1635075071 brs (0.000402%), ETA 489418866284y 68d 4h 41m 44s (elapsed 29m 54s)
			% nb. this quicksand one is actually the slower of 2 FABs, but so close in time to each other
			% that the one with substantially fewer interleavings. they were:
			% 145 interleavings; 1 preemptions; job time 23s; pldi time 475030016; new-fixed pldi cputime 3672836411
			% 225 interleavings; 2 preemptions; job time 32s; pldi time 465874053; new-fixed pldi cputime 3601564200
			% not like maximal found the bug at all so it's not like there's a comparison to make fair/unfair anyway
		{\tt lock\_fast}
		%({\tt spinlock})
			& 2,1	&\cpu{18.33}&\wtm{5.19}&\ints{2}&\cpu{*3.12}&\wtm{*3.12}&\ints{2}& \ETA{4} \\
			% 2 interleavings; 0 preemptions; job time 4s; pldi time 5191170; new-fixed pldi cputime 18328224
			% 2 interleavings; 0 preemptions; job time 2s; pldi time 3119081; new-fixed pldi cputime 3118912
		(perf)
			& 9,9	&\cpu{22.43}&\wtm{6.24}&\ints{2}&\cpu{*4.71}&\wtm{*4.71}&\ints{2}& \ETA{inf} \\
			% 2 interleavings; 0 preemptions; job time 3s; pldi time 6238515; new-fixed pldi cputime 22431755
			% 2 interleavings; 0 preemptions; job time 3s; pldi time 4712012; new-fixed pldi cputime 4711884
	\caption{Landslide's bug-finding performance on various test configurations.
		Iterative Deepening (\sect{\ref{sec:quicksand-id}}), optimized for fast bug-finding,
		is compared against Maximal State Space mode (\sect{\ref{sec:landslide-quicksand-options}}),
		optimized for fast verification.
		For each, I list the CPU-time and wall-clock time elapsed,
		plus the number of interleavings tested in the ultimately buggy state space
		until the bug was found.
		* marks the winning measurements between each series.
		Lastly, state space estimation (\sect{\ref{sec:landslide-estimate}})
		%though approximate at best,
		confers a sense of the exponential explosion.
	\label{tab:tm-buges}
\end{table*}

@@

In addition to the bugs I intentionally wrote in {\tt htm1} and {\tt swapbug},
%to my pleasant surprise % only the former of these is a surprise now
Landslide also found two bugs in the ``real-world'' transactional algorithms I tested.
\begin{itemize}
	\item
	{\bf Atomicity violation.}
	{\tt avl\_insert} with any parameters higher than (2,1)
	exposed a previously-unknown bug in the transactional AVL tree.
	Figure~\ref{fig:avlbug} shows the root cause, essentially the {\tt htm1} bug in disguise.
	This manifested
	alternately
	% landslide-trace-1533850116.527842.html
	% landslide-trace-1533850236.873032.html
	as a segfault (for test parameters (2,2) and (3,1))
	and
	% landslide-trace-1533854371.201589.html
	as a consistency-check assertion failure (for test parameters (2,3)).
	The presence of {\tt while (\_retry);} makes the necessary preemption window extremely small
	(between it and {\tt \_xbegin()}),
	making the bug extremely unlikely to manifest under stress testing,
	but Landslide is blind to such matters of chance.

@@

			& 2,2 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	&\ETAdag{36393y}&\ETAdag{35094477}\\
			% progress: 302022/35094477 brs (0.860597%), ETA 36392y 304d 6h 53m 45s (elapsed 9h 58m 49s)

@@

			& 2,2 & \ETAdag{qqq}	& \ETAdag{qqq}	&\cpu{16455.24}	& \ints{283756}	& \cpu{1285.06}	& \ints{21684} \\
			& 3,1 & \ETAdag{qqq}	& \ETAdag{qqq}	&\cpu{21153.64}	& \ints{366030}	& \cpu{705.04}	& \ints{12707} \\

@@

Firstly, the middle column of Table~\ref{tab:tm-verifs}
shows the impact of Landslide's experimental retry set reduction (\sect{\ref{sec:tm-retrysets}}).
Despite its conservative implementation,
it provides roughly 2-6x reduction in most tests, with up to 17x in extreme cases.
The biggest win is apparent in {\tt fig63},
the test contrived to induce as much reduction as possible
using transactional paths that conflict only with aborts and not with each other, and vice versa.
In {\tt fig63}(2,1), corresponding to Figure~\ref{fig:retry-set-example},
retry-set-enabled DPOR correctly prunes down to the optimal 6 interleavings,
while the baseline treats it identically to the fully-conflicting {\tt counter}(2,1).
Figure~\ref{fig:retry-reduction} depicts the difference between the state spaces explored by the two approaches.
